### 2025-11-24 paper/reviews/iclr2026/response.md

## Planned Revisions and New Experiments

### Improvements to RQ1 study

We are implementing the following improvements to the RQ1 dataset and evaluation:

- We are adding additional manual annotations to substantially increase the size of the hand-labelled subset. We will clarify that the subset was chosen by randomly choosing from each linguistic category.
- The SelectionAgent now also performs an initial labelling by linguistic category, in addition to identifying phrases to be replaced. We are evaluating this against the manual annotations to estimate the noise in the automated labelling.
- This allows us to derive an annotated dataset based on the full SciGen dataset, which we call SciGen-interpret. We will include this dataset as supplementary material, and release as an open source benchmark alongside the final version of the paper.
- We will redo the RQ1 study using the full dataset (adjusting confidence intervals for noise); this should remove some degenerate estimates (e.g. success rates of 0% and 100%) attributable to sparsity in the original subsample.
- We will perform an additional ablation study to assess the contribution of predefined helpers to success rates.

@@

For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot, where users iteratively refine the goal by nudging, re-prompting, and constraining via partial code. For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler agent-based architecture, where user interaction is restricted mainly to validation rather than iterative refinement.

# Response to Reviewer Bezq

> - Although the paper provides useful category-level statistics, it does not analyze statistical variance beyond reporting standard deviation or assess generalization to unseen writing styles or datasets.

We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we hope this addresses your point.

@@

Thank you for this point. A proper user study with the necessary experimental setup is out of scope for this work, but something we plan to do as the next step, alongside implementing a more complete interactive workflow. (We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.

> - The paper states that experiments were conducted on “a subsample of the SciGen dataset” (line 367), but it remains unclear how this subset was chosen. Could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?

For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues, resulting in 56 labelled replacement tasks for that dataset. We have since selected a substantially larger subsample (also randomly chosen) for manual annotation, which we will use for the final version of the paper.

In addition, we have added an automation pipeline to process every SciGen example using the SuggestionAgent. This means we can now evaluate the core InterpretationAgent on the *entirety* of the SciGen dataset (with confidence intervals adjusted to incorporate noise, estimated by evaluating SuggestionAgent annotations relative to the manual baseline).

@@

> - The system presented in this paper is more engineering-oriented, with relatively limited academic research value.
> - Overall, it can be regarded as a well-executed engineering paper.

We agree that it is an engineering-oriented paper, and appreciate your positive evaluation in that respect. However we also address important research questions around how an agent-based tool for retrofitting computational interpretations to scholarly articles might be designed, how current LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.

> - The experimental design lacks a comprehensive justification. Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered).

The lack of directly related prior work does indeed make comparison to baseline methods difficult. We agree more in-depth ablation studies are needed. We will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other similar experiments, and include those in our reporting on RQ1 and RQ2.

@@

# Response to Reviewer VRkd

> - My greatest confusion is: isn't this supposed to be an HCI paper? Shouldn't some version of this go to CHI/CSCW/UIST/...? To me, there isn't any technical contribution for an ML venue beyond getting the system implemented, but this type of contribution should be very suitable for the HCI venues.

The natural next step of this work will be a full-featured Copilot-like authoring (or reading) tool, which will enable a full user study. This will likely go to a venue like CHI, UIST, or IUI. For the present paper, our aim to validate key components of such a tool, including:

@@

- competency evaluation for LLMs (which we will substantially improve for this paper)
- proof-of-concept agentic architecture, with both automatic and manual validation (both important for this task)

We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper is in need further work. We are making improvements to that section as summarised in "Planned Revisions and New Experiments" above.

### 2025-11-25 paper/reviews/iclr2026/response.md

We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we hope this addresses your point.

In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the testing step, so all the SciGen examples were acting as out-of-distribution problems. We will make sure this is clear.

> - Can authors report statistics about the manual validation process? For examples, the ratio of accepted to rejected edits, average validation time per fragment, or the most common sources of rejection? Such data would illustrate how practical and scalable the workflow is for real authors.

Thank you for this point. A proper user study with the necessary experimental setup is out of scope for this work, but something we plan to do as the next step, alongside implementing a more complete interactive workflow. We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.
To illustrate how we plan to conduct our future user study, we briefly describe its basic design:
- Participants: scientific authors with experience in writing/reviewing manuscripts and knowledge of Fluid.
- Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them. The time taken for validation and the time for possible corrections will be measured. A comparison will be made with the baseline methodology in which participants will work without LLM support.
- Metrics: acceptance rate for the suggestions, average time for validation and corrections, and participant workload and satisfaction will be measured using standard questionnaires.
For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.

@@

- competency evaluation for LLMs (which we will substantially improve for this paper)
- proof-of-concept agentic architecture, with both automatic and manual validation (both important for this task)

We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper is in need further work. We are making improvements to that section as summarised in "Planned Revisions and New Experiments" above.

> - Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.

We agree that evaluating the tool with real users is important. However, this is beyond the scope of this work, as it would require a fully mature authoring tool. As discussed in more detail above (in the response to Reviewer Bezq), we plan to conduct a user study as future work, once the tool is finalised.

### 2025-11-25 paper/reviews/iclr2026/response.md

We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we hope this addresses your point.

In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the testing step, so all the SciGen examples were acting as out-of-distribution problems. We will make sure this is clear.

> - Can authors report statistics about the manual validation process? For examples, the ratio of accepted to rejected edits, average validation time per fragment, or the most common sources of rejection? Such data would illustrate how practical and scalable the workflow is for real authors.

Thank you for this point. A proper user study with the necessary experimental setup is out of scope for this work, but something we plan to do as the next step, alongside implementing a more complete interactive workflow. We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.

We anticipate a study with the following basic design:
- Participants: scientific authors with experience in writing/reviewing manuscripts and coding experience
- Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them. Time taken for validation and any corrections will be measured, and a comparison made with a baseline workflow in which participants work without LLM support.
- Metrics: acceptance rate for the suggestions, average time for validation and corrections, and participant workload and satisfaction will be measured using standard questionnaires.

For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.

### 2025-11-25 paper/reviews/iclr2026/response.md

- competency evaluation for LLMs (which we will substantially improve for this paper)
- proof-of-concept agentic architecture, with both automatic and manual validation (both important for this task)

We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper is in need further work. We are making improvements to that section as summarised in "Planned Revisions and New Experiments" above.

> - Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.

We agree that evaluating the tool with real users is important. However, this is beyond the scope of the present work, as it would require a mature authoring tool. As discussed in more detail above (in response to Reviewer Bezq), we plan to conduct a user study as future work once the tool is finalised.

### 2025-11-25 paper/reviews/iclr2026/response.md

We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we hope this addresses your point.

In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the testing step, so all the SciGen examples were acting as out-of-distribution problems. We will make sure this is clear.

> - Can authors report statistics about the manual validation process? For examples, the ratio of accepted to rejected edits, average validation time per fragment, or the most common sources of rejection? Such data would illustrate how practical and scalable the workflow is for real authors.

Thank you for this point. A proper user study with the necessary experimental setup is out of scope for this work, but we anticipate a study with the following basic design:
- Participants: scientific authors with experience in writing/reviewing manuscripts and coding experience
- Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them. Time taken for validation and any corrections will be measured, and a comparison made with a baseline workflow in which participants work without LLM support.
- Metrics: acceptance rate for the suggestions, average time for validation and corrections, and participant workload and satisfaction will be measured using standard questionnaires.

We will clarify in Future Work that this is something we plan to do as the next step, alongside implementing a more complete interactive workflow. (We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.

### 2025-11-25 paper/reviews/iclr2026/response.md

We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we hope this addresses your point.

In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the testing step, so all the SciGen examples were acting as out-of-distribution problems. We will make sure this is clear.

> - Can authors report statistics about the manual validation process? For examples, the ratio of accepted to rejected edits, average validation time per fragment, or the most common sources of rejection? Such data would illustrate how practical and scalable the workflow is for real authors.

Thank you for this point. A proper user study with the necessary experimental setup is out of scope for this work, but we anticipate a study with the following basic design:
- Participants: scientific authors with experience in writing/reviewing manuscripts and coding experience
- Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them; time taken for validation and any corrections will be measured, compared to a baseline workflow in which participants work without LLM support
- Metrics: acceptance rate for the suggestions, average time for validation and corrections, and participant workload and satisfaction will be measured using standard questionnaires.

We will clarify in Future Work that this is something we plan to do as the next step, alongside implementing a more complete interactive workflow. (We certainly recognise that there is a risk, as with AI-generated code in general, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
